---
title: "Assignment 2 - Task 1"
author: "Conner Smith"
date: "1/27/2022"
output: 
  html_document: 
    theme: flatly
    code_folding: hide
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
```

## **Overview**

This anslysis looks at the survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017. It uses binary logistic regression to predict whether a palmetto is species Serenoa repens or Sabal etonia. Data was collected by the Environemntal Data Initiative. 

## **Analysis**

This analysis uses binary logistic regression to test feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia. 

```{r}
# Read in the data 

palmetto <- read_csv(here("data", "all_palmetto_data.csv"))
```


### **Data Citation**

*Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5*
